---
title: "Drifter_GPS"
author: "Raine Detmer"
date: "4/6/2023"
output: html_document
---

README: code for analyzing Ohlmann et al (2019)'s drifter GPS data (Fig. S24 and S25)

These data are available at https://www.bco-dmo.org/dataset/739111 (metadata are also available at this link)

citation:
Ohlmann C (2019) Data from freely drifting kelp plants tagged with drifters in the Santa Barbara Channel between November of 2015 and December of 2017. Biological and Chemical Oceanography Data Management Office. https://doi.org/10.1575/1912/bco-dmo.739111.1



# packages

```{r}

library("tidyverse")

library("lubridate")

library("sf")

library("sp")

library("raster")#for working with rasters

library("mapview")#for making interactive maps

library("ggspatial")# for adding scale bar and north arrow to maps

```


# import data

```{r}
# Ohlmann et al (2019)'s data
driftgps <- read.csv("http://optserv1.whoi.edu/jg/listfnzm//dmoserv3.bco-dmo.org:80/BCO-DMO/Linking_Kelp_to_Beaches/kelp_drifter_data", header = F)

#View(driftgps) # has several header rows that need to be removed
#driftgps[4,] # this row needs to become the column names
driftgps <- driftgps[-c(1, 2, 3, 4),] %>% rename(id = V1, lat = V2, lon = V3, matime = V4, u = V5, v = V6, z = V7, date = V8, time = V9, ISO_DateTime_UTC = V10) %>% mutate(lat = as.numeric(lat), lon = as.numeric(lon), matime = as.numeric(matime), u = as.numeric(u), v = as.numeric(v), z = as.numeric(z))

#View(driftgps)

# The start and endpoints of individual deployments were only in the .mat file and not the .csv, so these were extracted from the .mat file in matlab and exported as "deplytimes.csv" (see "drifterGPS.m" in "intermediate_data_output/matlab formatting" folder for the code)
deploy <- read.csv("../../../intermediate_data_output/matlab formatting/deploytimes.csv", header = FALSE)
colnames(deploy) <- c("drifttrack", "id", "matimemin", "matimemax")

#View(deploy)
```

from the metadata:
id = drifter id
lat = latitude (decimal degrees)
lon = longitude (decimal degrees)
matime = Matlab datenum data type
u = u velocity (cm/s)
v = v velocity (cm/s)
z = water depth (m) interpolated from bathymetry data to each drifter position (neg. values are beneath the sea surface)
date = date in format yyyy-mm-dd
time = time in format HH:MM
ISO_DateTime_UTC = timestamp (UTC) in standard ISO 8601:2004(E) format YYYY-mm-ddTHH:MMZ

# format data

```{r}
#length(unique(driftgps$id)) # 20 drifters total (but were deployed multiple times)

# first format the dates/times
# make a new data set
driftgps2 <- driftgps

# add time3 column made by pasting the date and time
driftgps2$time3 <- paste(driftgps2$date, driftgps2$time, sep = " ")

# check conversion works
#as.POSIXct(driftgps2$time3[1], format="%Y-%m-%d %H:%M")

driftgps2$time4 <- as.POSIXct(driftgps2$time3, format="%Y-%m-%d %H:%M")
#View(driftgps2)

# now select just the id, lat, lon, matimes, velocities, depth, date, and this new date time object
driftgps3 <- driftgps2[, c(1:3, 4:8, 12)]

#View(driftgps3)

# join with the data on drift deployments (individual tracks for each drifter)
# need to round the matimes in order to match them up (one dataset has higher precision than the other)
# round matimes to match number of decimal places in driftgps3
deploy2 <- deploy %>% mutate(matimemin = round(matimemin, 4), matimemax = round(matimemax,4)) %>% mutate(id = as.character(id))
#View(deploy2)

# join the full gps data with the data on individual tracks for each drifter, filter out the rows where the times in the full data set match the start or end times of the tracks
driftgps4 <- left_join(driftgps3, deploy2, by = c("id")) %>% filter(matime==matimemin | matime==matimemax)

# check this
#dim(driftgps4) # 297 entries
#length(deploy2$drifttrack)*2 # 294 (147 deployments, each with min and max time)

# which have more than 2?
#driftgps4 %>% group_by(drifttrack, id) %>% summarize(nrows = n()) %>% filter(nrows !=2)
#View(driftgps4[which(driftgps4$id==37 & driftgps4$drifttrack %in% c(4, 5)),])
#View(driftgps3[which(driftgps3$id=="37"),])

# rows 41 and 42 each have id= 37 with the same position; seems like it was in 2 places at once
# row 45-47 have 3 more 37's for the same two drift tracks (4 and 5) as above
# so there are 7 id 37's total for tracks 4 and 5 when there should only be 4

#View(deploy2[which(deploy2$id=="37"),]) # the tracks 4 and 5 have the same end point...

# from the Matlab data: end matime points for tracks 4 and 5
#7.363547014004600e+05
#7.363547014004600e+05 # end points are the same in the .mat data too

# there appear to be two sets of duplicates: one for track 4 and one for track 5 (which seems to overlap with track 4- has later start point but same end point)

# first remove all the track 5 rows for id 37
#which(driftgps4$id==37 & driftgps4$drifttrack==5)
driftgps5 <- driftgps4[-c(40, 42, 45, 47), ]

# now just choose the first of the duplicate track 4 rows
#which(driftgps5$id==37 & driftgps5$latitude==34.403534 & driftgps5$longitude==-119.863446)
driftgps5 <- driftgps5[-43, ]

# now join driftgps5 back with driftgps3 by id and matime and fill in the drift track column across all gps points for each id's track
driftgps6 <- left_join(driftgps3, driftgps5, by = c("id", "lat", "lon", "matime", "u","v","z","date", "time4")) %>% fill(drifttrack, .direction = "down") %>% select(-matimemin, -matimemax)

#View(driftgps6)

```


turn the dataset into spatial object

```{r}

# turn the gps data into an sf object, where the lat/lon positions are sf_points
#convert the dataframe to sf object
driftpts <- st_as_sf(driftgps6, coords = c("lon", "lat"))

st_crs(driftpts) <- 4326#set the crs

#driftpts

#convert to utm (to match the patch data)
driftpts2 <- st_transform(driftpts, 32611)#change the coordinate reference system to utm zone 11

#add columns with the coordinate values
driftpts2$utmx <- st_coordinates(driftpts2)[,1]
driftpts2$utmy <- st_coordinates(driftpts2)[,2]

```

```{r}
# look at some of the tracks
mapview(driftpts[which(driftpts$id==31 & driftpts$drifttrack==1),])#+ mapview(patchmerge)

mapview(driftpts[which(driftpts$id==38 & driftpts$drifttrack==1),]) #+ mapview(patchmerge)

mapview(driftpts[which(driftpts$id==40 & driftpts$drifttrack==1),]) #+ mapview(patchmerge)

# this one is really long
mapview(driftpts[which(driftpts$id==41 & driftpts$drifttrack==19),])

```


# patch data

get the data on metapopulation patch locations and convert to a raster

```{r}

#download data on patch coordinates

#https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sbc.101.2

# Package ID: knb-lter-sbc.101.2 Cataloging System:https://pasta.edirepository.org.
# Data set title: SBC LTER: Spatial definitions of giant kelp (Macrocystis pyrifera) patches in southern and central California.
# Data set creator:  Kyle C Cavanaugh -  
# Data set creator:  David A Siegel -  
# Data set creator:  Peter T Raimondi -  
# Data set creator:  Filipe Alberto -  
# Contact:    - Information Manager, Santa Barbara Coastal LTER   - sbclter@msi.ucsb.edu
# Stylesheet v2.11 for metadata conversion into program: John H. Porter, Univ. Virginia, jporter@virginia.edu 

patchdef <-read.csv("https://pasta.lternet.edu/package/data/eml/knb-lter-sbc/101/2/0d416cc00ad948c6befb2a97e9db6414", header=F, skip=1, sep="," , col.names=c("patch_number","pixel_latitude", "pixel_longitude"), check.names=TRUE)


```

```{r}

#View(patchdef)#need to rename the lat and long columns and add a pixel_id column (each row is a unique Landsat pixel)

patchdef2 <- patchdef %>% rename(latitude = pixel_latitude, longitude = pixel_longitude)%>% mutate(pixid = row_number())#add pixel id number for comparing before and after shift

patchdef2$patch_number <- gsub("socal_", "", patchdef$patch_number) #the patches are labeled as "socal_#" so extract just the number

#now turn this into a spatial object, change the coordinate system to utm, and change the pixel coordinates from the center to the upper left (for rasterizing the points)

#convert the dataframe to sf object
patchpts <- st_as_sf(patchdef2, coords = c("longitude", "latitude"))
st_crs(patchpts) <- 4326#set the crs

#convert to utm
patchpts2 <- st_transform(patchpts, 32611)#change the spatial reference system to utm zone 11

#add columns with the coordinate values
patchpts2$utmx <- st_coordinates(patchpts2)[,1]
patchpts2$utmy <- st_coordinates(patchpts2)[,2]

#change the coordinates from the pixel centers to the upper left corners: subtracting from the x and adding to the y to go from center to upper right
patchpts2$newx <- round(patchpts2$utmx) - 15
patchpts2$newy <- round(patchpts2$utmy) +15
#Reason for rounding: when converting from degree decimal to utm, the utm coords had a lot of decimal places, and the distances between adjacent points wasn't exactly 30 (it was around 29.999). And this caused some points to be lost when turning them into a raster (which requires and evenly spaced grid) 

#now make a new sf object with the upper left points as the coordinates
#first make this a data frame
patchpts2.df <- as.data.frame(patchpts2)#make this a data frame

#remove unnecessary rows and change the coordinates, also change patch_number to patch
patchpts2.df <- patchpts2.df %>% dplyr::select(patch_number, newx, newy, pixid) %>% relocate(pixid, .before = patch_number) %>% rename(patch = patch_number)

#make this new data frame an sf object using the new coordinates for the point geometries 
patchpts3 <- patchpts2.df %>% st_as_sf(coords = c("newx", "newy")) %>% st_set_crs(32611)

#add utmx and utmy columns
patchpts3$utmx <- st_coordinates(patchpts3)[,1]
patchpts3$utmy <- st_coordinates(patchpts3)[,2]
 
#check that it worked
#mapview(list(patchpts2[1:500, ],patchpts3[1:500, ]), col.regions=list("red","blue"), col=list("red","blue")) 

```


Filter out study region and turn the points into a raster (each point represents a 30 x 30m Landsat pixel)

```{r}
#Filter out study region (which is approximately bordered by xmin: 168810 ymin: 3703890 xmax: 327120 ymax: 3829740)
patchpts_sub <- patchpts3

patchpts_sub$utmx <- st_coordinates(patchpts_sub)[,1]
patchpts_sub$utmy <- st_coordinates(patchpts_sub)[,2]

#subset out the study region
patchpts_sub1 <- patchpts_sub %>% filter(utmx<327120)%>% filter(utmx>168810) %>% filter(utmy>3703890) %>% filter(utmy < 3829740)

patchpts_sub1$patch <- as.numeric(patchpts_sub1$patch)#needed to change patch to numeric for rasterize


#make the raster

#get the max and min values of the raster: the coords are the upper left of each cell. So to get the bottom and right borders, need to add 30m to max lon and subtract 30m from the min lat
pxmnutm <- min(patchpts_sub1$utmx) 
pxmxutm <- max(patchpts_sub1$utmx) + 30
pymnutm <- min(patchpts_sub1$utmy) - 30
pymxutm <- max(patchpts_sub1$utmy) 


#make the raster with same crs as the data points, no values (empty), resolution of 30m x 30m, and the min and max values specified above
patch_rst <- raster(crs = crs(patchpts_sub1), vals = 0, resolution = c(30, 30), xmn=pxmnutm, ymn=pymnutm, xmx=pxmxutm, ymx=pymxutm)

#area(patch_rst)#check that the area of the cells is 900m2

#rasterize the points
patch_rst2 <- rasterize(patchpts_sub1, patch_rst)

#now convert this back to sf object with the cells as polygons
#https://taromieno.netlify.app/post/raster_to_polygons_2018/

#first convert to sp object (takes a few min)
patchsp <- as(patch_rst2,'SpatialPolygonsDataFrame')

#check the number of features (polygons) match number of data points
#patchpts_sub1#112467
#patchsp #112467

#now convert this to sf object
patchrstsf <- st_as_sf(patchsp)%>% st_set_crs(32611)

#check that it looks good: plot the polygons and the points
#mapview(patchrstsf[which(patchrstsf$patch==160), ])+ patchpts_sub1[which(patchpts_sub1$patch==160), ] 

#mapview(patchrstsf[which(patchrstsf$patch==160), ], col.regions=list("red"),col=list("red"))+ mapview(list(patchpts3[which(patchpts3$patch==160), ], patchpts2[which(patchpts2$patch_number==160), ]), col.regions=list("red","blue"),col=list("red","blue")) 


# now merge all the patch pixels into a single polygon for each patch
patchmerge <- patchrstsf %>% 
    group_by(patch) %>%
    summarise(geometry = sf::st_union(geometry)) %>%
    ungroup()


#mapview(patchmerge)

```


# intersection

get the intersections of the drifter tracks and metapopulation patches

```{r}

# want the intersection of driftpts2 and patchmerge
patch_tracks <- st_intersects(driftpts2, patchmerge)

#patch_tracks # length = number of gps points, corresponding number is the row of patchmerge that that gps point intersects with. e.g, first row has element 51, which means the gps point in row 1 of driftpts2 was in the patch in row 51 of patchmerge. Empty rows means the gps point in that row wasn't in any patch

# turn patch_tracks into a data frame with the gps point and patchmerge rows as columns (note this automatically gets rid of the empty rows = rows where gps point wasn't in any patch)
patch_tracks_match <- as.data.frame(patch_tracks)
colnames(patch_tracks_match) <- c("gpsrow", "patchrow")

#length(unique(patch_tracks_match$patchrow)) # 20 patches total 

# can map everything at once to make sure it worked, but this takes a while to load (there are a lot of points)
#mapview(driftpts2[c(patch_tracks_match$gpsrow), ]) + mapview(patchmerge[c(patch_tracks_match$patchrow),])


# patch_tracks_match just gives the row numbers in driftpts and patchmerge, so need to get the actual drift point information and the patch number
patch_tracks_match2 <- patch_tracks_match

# make holding columns
patch_tracks_match2$id <- NaN*patch_tracks_match2$gpsrow # drifter id

patch_tracks_match2$drifttrack <- NaN*patch_tracks_match2$gpsrow # the track

patch_tracks_match2$matime <- NaN*patch_tracks_match2$gpsrow # drifter timepoint

patch_tracks_match2$patch <- NaN*patch_tracks_match2$gpsrow # patch

for(i in 1:length(patch_tracks_match2$gpsrow)){ # for each row in patch_tracks_match2

patch_tracks_match2$id[i] <- driftpts$id[patch_tracks_match2$gpsrow[i]] # drifter id = id in the row of driftpts given by the ith value of patch_tracks_match2$gpsrow

patch_tracks_match2$drifttrack[i] <- driftpts$drifttrack[patch_tracks_match2$gpsrow[i]] # the track

patch_tracks_match2$matime[i] <-driftpts$matime[patch_tracks_match2$gpsrow[i]] # drifter timepoint

patch_tracks_match2$patch[i] <- patchmerge$patch[patch_tracks_match2$patchrow[i]] # patch = patch in the row of patchmerge given by the ith value of patch_tracks_match2$patchrow
  
}


#View(patch_tracks_match2)

#remove the index columns
#patch_tracks_match2 <- patch_tracks_match2 %>% dplyr::select(pixel, patch, site.trans)


# remove the row indexing columns and join this with the rest of the information on the drifters
#patch_tracks_match3 <- patch_tracks_match2 %>% dplyr::select(-gpsrow, -patchrow) %>% left_join(driftgps6, by = c("id", "matime","drifttrack"))

#View(patch_tracks_match3)



```

## Fig S24

plot an example trajectory

```{r}


# get the base map
# download the usa map data
usa <- raster::getData("GADM", country = c("United States"), level = 1)


# get the boundary region (from boundaries of socal metapop: https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.101)
bounds_full <- st_polygon(list(rbind(c(-120.754, 35.178),# upper left corner
                                       c(-117.129, 35.178), # upper right corner
                                       c(-117.129, 32.504), # lower right corner
                                       c(-120.754, 32.504),  # lower left corner, then complete the polygon
                                       c(-120.754, 35.178)))) %>% st_sfc() %>% st_set_crs(4326)# st_sfc turns this into a simple feature geometry list column, and then set the coordinate system with st_set_crs

# now to take the usa data, turn it into an sf object, get the part the overlaps with the boundary region, and then set the coordinate system, and then st_union makes the whole thing a single multipolygon
# note this takes a few min  
shore_full <- usa %>% sf::st_as_sf() %>% sf::st_intersection(bounds_full) %>% sf::st_transform(4326) %>% sf::st_union()

# check trajectories
#mapview(driftpts[which(driftpts$id==41 & driftpts$drifttrack==5),])+ mapview(patchmerge)

#mapview(driftpts[which(driftpts$id==50 & driftpts$drifttrack==13),])+ mapview(patchmerge)

```


```{r}
pdf("driftmap.pdf")
ggplot() +
    geom_sf(data = shore_full, fill = "gray80", color = NA)+# fill = color to make the land, color = NA removes the border around the land
  geom_sf(data = patchmerge, color = "#78A0A0", fill = "#78A0A0", linewidth = 1)+#goldenrod2
  geom_sf(data = driftpts[which(driftpts$id==41 & driftpts$drifttrack==5),], color = "deepskyblue", fill = "deepskyblue", size = 1)+#shape = 23 makes diamond
  #geom_sf_text(data = citysf2, aes(label = city),color = "black", size = 3)+
  annotate(geom = "text", x = -119.8489, y = 34.4140, size = 3.5, label = "UCSB", color = "black")+#fontface = "bold"
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill = 'gray96', color = 'gray96'), axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), axis.title = element_blank())+# remove gridlines, change background (=ocean) color, change size of x and y axis labels, remove axis titles
  scale_y_discrete(breaks = c(34.4, 34.44))+#specify xaxis labels
  scale_x_discrete(breaks = c(-119.9, -119.8, -119.7))+#specify xaxis labels
  annotation_scale(location = "bl", style = "ticks",  pad_x = unit(0.25, "cm"), pad_y = unit(0.45, "cm"), text_cex=0.9)+
  #annotation_north_arrow(location = "bl", style = north_arrow_orienteering(text_size = 8), height = unit(0.5, "cm"), width = unit(0.5, "cm"), pad_x = unit(0.25, "cm"), pad_y = unit(1, "cm"))+# change the padding to adjust north arrow location in the specified corner
  annotation_north_arrow(location = "tr", style = north_arrow_orienteering(text_size = 11), height = unit(0.7, "cm"), width = unit(0.7, "cm"))+# change the padding to adjust north arrow location in the specified corner
  #coord_sf(xlim = c(-120.754, -117.129), ylim = c(32.504, 35.178), expand = F)# make the map go all the way to the edges of the plot
  coord_sf(xlim = c(-119.95, -119.65), ylim = c(34.38, 34.45), expand = F)# make the map go all the way to the edges of the plot
#coord_sf(xlim = c(-120.6093 , -117.2381), ylim = c(32.64389, 34.55651), expand = F)# make the map go all the way to the edges of the plot
dev.off()

```

# paths through patches

look at the percent of drifters that started in one patch that ended up in/went through another patch (to get an idea of how frequently kelp passes through patches other than the one it started in)

```{r}
# use the full drifter data, join with the patch data, so all drifter locations in a patch have the patch number and all locations not in a patch have NA, then group by drift id and track and get the number of patches encountered

# make a data frame with all the gps data and also whether each gps location was in a patch, and which patch it was in
# start by copying the drift gps data
driftgps_patch <- driftgps6
# add a column with the row id
driftgps_patch$gpsrow <- seq(from = 1, to = length(driftgps_patch$id), by = 1)
# join the patch_tracks_match data to the driftgps data, but this time keep all the drift gps points
driftgps_patch <- left_join(driftgps_patch, patch_tracks_match2, by = c("id","drifttrack", "gpsrow", "matime"))%>% mutate(id.track = paste(id, drifttrack, sep = ".")) # add a id.track column (each unique combination of drifter id and its deployment track)

# check number of rows match
#length(driftgps_patch$id)
#length(driftgps6$id)

#View(driftgps_patch)

# what fraction of observations were in patches?
#length(driftgps_patch$gpsrow[which(is.na(driftgps_patch$patch)==F)])/length(driftgps_patch$gpsrow) # 43% 

# now get the id.tracks that started in a patch
# for each id.track, get the row corresponding to the first transmission
driftgps_start <- driftgps_patch %>% group_by(id, drifttrack, id.track) %>% filter(time4 == min(time4)) 

length(driftgps_start$id.track) # should be the same as the number of unique id.track values in the full data
length(unique(driftgps_patch$id.track)) # same (146)

length(driftgps_start$id.track[which(is.na(driftgps_start$patch)==F)])/length(driftgps_start$id.track) # about 20% started in patches

#length(which(is.na(driftgps_start$patch)==F)) #29/146 started in a patch

# get the id.tracks where the drifter started in a patch
start_patch <- driftgps_start$id.track[which(is.na(driftgps_start$patch)==F)]

# now want to get the number of patches (excluding starting patches) that each drifter went through
# add patch indicator column with 1 (in patch) or 0 (not in a patch)
driftgps_patch2 <- driftgps_patch %>% group_by(id, drifttrack, id.track, patch) %>% summarize(matime = mean(matime)) %>% ungroup()%>% mutate(Ipatch = if_else(is.na(patch)==F, 1, 0))%>% group_by(id, drifttrack, id.track) %>% summarise(n_patches = sum(Ipatch)) %>% mutate(n_patches2 = if_else(id.track %in% start_patch, n_patches-1, if_else(id.track %in% start_patch))) 

#View(driftgps_patch2)


length(driftgps_patch2$id.track[which(driftgps_patch2$n_patches>0)])/length(unique(driftgps_patch$id.track))# 95% of the drifter tracks went through at least one patch 

length(driftgps_patch2$id.track[which(driftgps_patch2$n_patches2>0)])/length(unique(driftgps_patch$id.track))# 85% of the drifter tracks went through at least one patch other than the starting patch (or didn't start in a patch and went through at least one patch)

length(driftgps_patch2$id.track[which(driftgps_patch2$n_patches>1)])/length(unique(driftgps_patch$id.track))# 56% of the drifter tracks went through at least two patches



# 57.2 went through 6 patches
#mapview(driftpts[which(driftpts$id==57 & driftpts$drifttrack==2),]) + mapview(patchmerge)

# check what time of year this one
#driftpts[which(driftpts$id==57 & driftpts$drifttrack==2),] # this was in December 2015, so might have been low kelp in the patches (less likely to get tangled up)


```

## Fig S25
```{r}

pdf("driftpatchhist.pdf", width = 10)
driftgps_patch2 %>% 
  ggplot(aes(as.factor(n_patches))) + 
  geom_bar()+
  xlab("Number of patches encountered")+
  ylab("Number of drifters")+
  theme_bw() + #make the background white not gray
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_blank(), axis.line = element_line(), axis.text=element_text(size=14),
        axis.title=element_text(size=18)) #remove gridlines
dev.off()

```

# mean distance

get the distance between the first and last patches enountered by each drifter track

first get the first and last patches encountered by each plant that went through at least 2 patches
```{r}

# get the id.tracks for which a plant went through at least 2 patches
btwpatch <- driftgps_patch2$id.track[which(driftgps_patch2$n_patches>1)]


#first get the first and last patches encountered
driftgps_patchmn <- driftgps_patch %>% filter(id.track %in% btwpatch) %>% dplyr::select(id, matime, drifttrack, patch, id.track) %>% filter(is.na(patch)==F)%>% group_by(id, drifttrack, id.track) %>% mutate(mintime = min(matime)) %>% ungroup() %>% filter(matime== mintime) %>% rename(patchmn = patch)

#View(driftgps_patchmn)

driftgps_patchmxmn <- driftgps_patch %>% filter(id.track %in% btwpatch)%>% dplyr::select(id, matime, drifttrack, patch, id.track) %>% filter(is.na(patch)==F)%>% group_by(id, drifttrack, id.track) %>% mutate(maxtime = max(matime)) %>% ungroup() %>% filter(matime==maxtime)%>% rename(patchmx = patch) %>% left_join(driftgps_patchmn, by = c("id", "drifttrack", "id.track")) %>% mutate(patchstartstop = paste(patchmn, patchmx, sep = "."))

#View(driftgps_patchmxmn)

#head(driftgps_patch)

#mapview(driftpts[which(driftpts$id==38 & driftpts$drifttrack==18),]) + mapview(patchmerge)

#mapview(driftpts[which(driftpts$id==60 & driftpts$drifttrack==13),]) + mapview(patchmerge)


```

now get the distances between patches

the coordinates of the centers of the patches are available from 

https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.102

```{r}
# Package ID: knb-lter-sbc.102.1 Cataloging System:https://pasta.lternet.edu.
# Data set title: Kelp metapopulations: Semi-annual time series of giant kelp patch area, biomass and fecundity in southern California, 1996 - 2006.
# Data set creator:  Max Castorani -  
# Data set creator:  Daniel Reed -  
# Data set creator:  Peter Raimondi -  
# Data set creator:  Filipe Alberto -  
# Data set creator:  Tom Bell -  
# Data set creator:  Kyle Cavanaugh -  
# Data set creator:  David Siegel -  
# Contact:    - Information Manager, Santa Barbara Coastal LTER   - sbclter@msi.ucsb.edu
# Stylesheet v2.11 for metadata conversion into program: John H. Porter, Univ. Virginia, jporter@virginia.edu 
# Metadata Link: https://portal.lternet.edu/nis/metadataviewer?packageid=knb-lter-sbc.102.1


patchUrl  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-sbc/102/1/fdeb932980156bbc1dd56622619dc2dd" 
patchFile <- tempfile()
try(download.file(patchUrl,patchFile,method="curl"))
if (is.na(file.size(patchFile))) download.file(patchUrl,patchFile,method="auto")

                   
 patch_stats <-read.csv(patchFile,header=F,skip=1,sep=",",quot='"', col.names=c("patch_number", "pixel_latitude", "pixel_longitude","patch_area","year", "semester", "patch_biomass","patch_fecundity"), check.names=TRUE)
               
unlink(patchFile)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                
if (class(patch_stats$patch_number)!="factor") patch_stats$patch_number<- as.factor(patch_stats$patch_number)
if (class(patch_stats$pixel_latitude)=="factor") patch_stats$pixel_latitude <-as.numeric(levels(patch_stats$pixel_latitude))[as.integer(patch_stats$pixel_latitude) ]               
if (class(patch_stats$pixel_latitude)=="character") patch_stats$pixel_latitude <-as.numeric(patch_stats$pixel_latitude)
if (class(patch_stats$pixel_longitude)=="factor") patch_stats$pixel_longitude <-as.numeric(levels(patch_stats$pixel_longitude))[as.integer(patch_stats$pixel_longitude) ]               
if (class(patch_stats$pixel_longitude)=="character") patch_stats$pixel_longitude <-as.numeric(patch_stats$pixel_longitude)
if (class(patch_stats$patch_area)=="factor") patch_stats$patch_area <-as.numeric(levels(patch_stats$patch_area))[as.integer(patch_stats$patch_area) ]               
if (class(patch_stats$patch_area)=="character") patch_stats$patch_area <-as.numeric(patch_stats$patch_area)
if (class(patch_stats$semester)!="factor") patch_stats$semester<- as.factor(patch_stats$semester)
if (class(patch_stats$patch_biomass)=="factor") patch_stats$patch_biomass <-as.numeric(levels(patch_stats$patch_biomass))[as.integer(patch_stats$patch_biomass) ]               
if (class(patch_stats$patch_biomass)=="character") patch_stats$patch_biomass <-as.numeric(patch_stats$patch_biomass)
if (class(patch_stats$patch_fecundity)=="factor") patch_stats$patch_fecundity <-as.numeric(levels(patch_stats$patch_fecundity))[as.integer(patch_stats$patch_fecundity) ]               
if (class(patch_stats$patch_fecundity)=="character") patch_stats$patch_fecundity <-as.numeric(patch_stats$patch_fecundity)
                
```

```{r}
# get the list of patches that intersected with drifter tracks
focalpatches <- unique(c(driftgps_patchmxmn$patchmn, driftgps_patchmxmn$patchmx))

#filter out the patches in the drifter data and turn the patch coords into sf object (also need distinct because these data have observations from multiple time points but patch location doesn't change)
patch_stats2 <- patch_stats %>% rename(patch = patch_number, lon = pixel_longitude, lat = pixel_latitude) %>% filter(patch %in% focalpatches) %>% dplyr::select(patch, lon, lat) %>% distinct()%>% arrange(patch) %>% mutate(patchid = row_number())# order by patch number

#View(patch_stats2)

#patchdat2$patch_number

# turn into sf object
patch_statssf <-patch_stats2 %>% st_as_sf(coords = c("lon","lat")) %>% st_set_crs(4326)

#mapview(patchmerge) + mapview(patch_statssf)

#get distance matrix with all pairwise distances 
patchdist <- st_distance(patch_statssf)

dim(patchdist)#19 x 19

#patchdist

#as.numeric(patchdist)# turns this into a vector, where entries 1:19 = distance between patch 1 and patch 1:19, entries 20: 20+19 = distance between patch 2 and patches 1:19, etc.

# can use expand.grid:
# expand.grid(c(1, 2, 3), c(1, 2, 3))
patch.grid <- as.data.frame(expand.grid(c(1:19), c(1:19)))

#View(patch.grid)


# turn into data frame
patchdist.df <- as.data.frame(as.numeric(patchdist))
colnames(patchdist.df) <- c("distance")

#View(patchdist.df)

patchdist.df$start_patch <- patch.grid[ ,2]
patchdist.df$end_patch <- patch.grid[,1]

#View(patchdist.df)

# convert the patch indexes into the actual patches
patch_stats_start <- patch_stats2 %>% rename(start_patch = patchid)
patch_stats_end <- patch_stats2 %>% rename(end_patch = patchid)

# change distance from m to km
patchdist.df2 <- patchdist.df %>% mutate(distance = distance/1000) %>% left_join(patch_stats_start, by = "start_patch") %>% rename(startingpatch = patch) %>% dplyr::select(-start_patch, -end_patch, -lat, -lon)

patchdist.df3 <- patchdist.df %>% mutate(distance = distance/1000) %>% left_join(patch_stats_end, by = "end_patch") %>% rename(endingpatch = patch) %>% dplyr::select(-end_patch, -start_patch, -lat, -lon, -distance)

patchdist.df4 <- cbind(patchdist.df2, patchdist.df3) %>% mutate(patchstartstop = paste(startingpatch, endingpatch, sep = ".")) %>% right_join(driftgps_patchmxmn, by = c("patchstartstop"))

#View(patchdist.df4)

```


```{r}
#hist(patchdist.df4$distance, n = 20)

mean(patchdist.df4$distance) # 3.5

# without the large outlier
mean(patchdist.df4$distance[which(patchdist.df4$distance<60)]) # 2.8

sd(patchdist.df4$distance)

```







