---
title: "ODE_random_effects"
author: "Raine Detmer"
date: "4/7/2023"
output: html_document
---

README: getting the values/distributions of the ODE random effect parameters (and associated supplemental figures S21 and S22)


# packages

```{r}
library("tidyverse")

library("lubridate")

```

# data

import the already-formatted data sets used to fit the GLMM (formatted in "format_data.Rmd")

```{r}
# benthic transect data
benthicdt <- read.csv("../../../intermediate_data_output/GLMM_data/benthic_data.csv") %>% dplyr::select(-X)

# metapopulation patch data
patchdt <- read.csv("../../../intermediate_data_output/GLMM_data/patch_data.csv") %>% dplyr::select(-X)

# convert patch connectivity to 1 semester lag
patchdt1Sfull <- patchdt %>% group_by(patch) %>% mutate_at(c(colnames(patchdt[4:9])), list(~ lag(.,n=1L, default = NA, order_by = year.sem))) %>% ungroup()

# join the benthic transect and patch connectivity data
full_datalvals <-left_join(benthicdt, patchdt1Sfull, by = c("year", "semester", "year.sem","patch")) %>% filter(is.na(patch_area)==F) %>% mutate(patch = as.factor(patch), transect_id = as.factor(transect_id))%>% filter(is.na(patch_conn0.9)==F, is.na(kelp_dens)==F)

```

## plant biomass
import SBC LTER and KFM data on kelp plant biomass

```{r}
# LTER data: https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.18

# Package ID: knb-lter-sbc.18.26 Cataloging System:https://pasta.edirepository.org.
# Data set title: SBC LTER: Reef: Kelp Forest Community Dynamics: Abundance and size of Giant Kelp (Macrocystis Pyrifera), ongoing since 2000.
# Data set creator:    - Santa Barbara Coastal LTER 
# Data set creator:  Daniel C Reed -  
# Data set creator:  Robert J Miller -  
# Contact:    - Information Manager, Santa Barbara Coastal LTER   - sbclter@msi.ucsb.edu
# Stylesheet v2.11 for metadata conversion into program: John H. Porter, Univ. Virginia, jporter@virginia.edu 

 lter_kelp <-read.csv("https://pasta.lternet.edu/package/data/eml/knb-lter-sbc/18/26/e0144d254ab3b45fa1cd9d0a0420e2e8",header=F, skip=1,sep=",",quot='"', col.names=c("YEAR","MONTH","DATE","SITE","TRANSECT","QUAD", "SIDE","SP_CODE","FRONDS","HLD_DIAM","AREA","OBS_CODE","NOTES","SCIENTIFIC_NAME","COMMON_NAME","TAXON_KINGDOM","TAXON_PHYLUM","TAXON_CLASS","TAXON_ORDER","TAXON_FAMILY","TAXON_GENUS","GROUP","SURVEY", "MOBILITY","GROWTH_MORPH"), check.names=TRUE)

# replace NAs
 lter_kelp <-  lter_kelp %>% mutate_at(c(colnames(lter_kelp[9:11])), list(~if_else(.==-99999, as.integer(NA), .)))
 
 
 # KFM giant kelp size-frequency data from Kushner et al. (2013): https://esapubs.org/archive/ecol/E094/245/metadata.php#_Hlk327197899
 kfm_kelp <- read.csv("https://esapubs.org/archive/ecol/E094/245/Giant%20kelp%20size%20frequency.csv")
 
 # regression pars to convert fronds to biomass (from SBC LTER NPP data: https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.112)
 # Package ID: knb-lter-sbc.112.7 Cataloging System:https://pasta.edirepository.org.
# Data set title: SBC LTER: REEF:  Net primary production, growth and standing crop of Macrocystis pyrifera in Southern California.
# Data set creator:    - Santa Barbara Coastal LTER 
# Data set creator:  Andrew A Rassweiler -  
# Data set creator:  Shannon Harrer -  
# Data set creator:  Daniel C Reed -  
# Data set creator:  Clint J Nelson -  
# Data set creator:  Robert J Miller -  
# Contact:    - Information Manager, Santa Barbara Coastal LTER   - sbclter@msi.ucsb.edu
# Stylesheet v2.11 for metadata conversion into program: John H. Porter, Univ. Virginia, jporter@virginia.edu 
 kelp_regpars <-read.csv("https://pasta.lternet.edu/package/data/eml/knb-lter-sbc/112/5/6bc968141edf7b9e0a6bac923f5f78b2", header=F,skip=1,sep="," ,quot='"', col.names=c("Model", "Month", "Independent_variable", "Dependent_variable", "Slope", "Intercept", "pvalue", "rsquare","N"), check.names=TRUE)
 
 #subset out model 1 (indep variable = fronds/m2 in the specified month, dep variable = kg dry mass/m2 in the specified month)
kelp_regpars <- kelp_regpars %>% filter(Model==1) %>% dplyr::select(Month, Independent_variable, Dependent_variable, Slope, Intercept) %>% rename(MONTH = Month) #%>% mutate(MONTH = as.factor(MONTH))

```


## juvenile densities
import the SBC LTER and KFM data on juvenile kelp densities

```{r}
# LTER data: https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.19

# Package ID: knb-lter-sbc.19.28 Cataloging System:https://pasta.edirepository.org.
# Data set title: SBC LTER: Reef: Kelp Forest Community Dynamics: Invertebrate and algal density.
# Data set creator:    - Santa Barbara Coastal LTER 
# Data set creator:  Daniel C Reed -  
# Data set creator:  Robert J Miller -  
# Contact:    - Information Manager, Santa Barbara Coastal LTER   - sbclter@msi.ucsb.edu
# Stylesheet v2.11 for metadata conversion into program: John H. Porter, Univ. Virginia, jporter@virginia.edu 

 jkelpLTER <-read.csv("https://pasta.lternet.edu/package/data/eml/knb-lter-sbc/19/28/5daf0da45925ba9014872c6bc9f6c8bb", header=F,skip=1 ,sep=",",quot='"', col.names=c("YEAR","MONTH","DATE", "SITE","TRANSECT","QUAD","SIDE","SP_CODE","SIZE", "COUNT","AREA", "SCIENTIFIC_NAME","COMMON_NAME", "TAXON_KINGDOM","TAXON_PHYLUM","TAXON_CLASS","TAXON_ORDER","TAXON_FAMILY", "TAXON_GENUS","GROUP","SURVEY","MOBILITY","GROWTH_MORPH"), check.names=TRUE)
               
# replace NAs
 jkelpLTER <-  jkelpLTER %>% mutate(COUNT = as.numeric(COUNT), AREA = as.numeric(AREA)) %>% mutate_at(c(colnames(jkelpLTER[9:10])), list(~if_else(.==-99999, as.numeric(NA), .)))
 
 jkelpLTER <- jkelpLTER %>% filter(SCIENTIFIC_NAME == "Macrocystis pyrifera")
 
 # use factor() to drop factor levels in the full data but not in the subsetted data (see https://www.edureka.co/community/1593/how-to-drop-factor-levels-in-a-subsetted-data-frame)
jkelpLTER[] <- lapply(jkelpLTER, function(x) if(is.factor(x)) factor(x) else x)

#levels(as.factor(jkelpLTER$SP_CODE))# all are juveniles (MPJ)
 
# filter out just juvenile kelp
 
# KFM data: benthic density data from Kushner et al. (2013) 
# https://esapubs.org/archive/ecol/E094/245/metadata.php#_Hlk327197722

 jkelpKFM <- read.csv("https://esapubs.org/archive/ecol/E094/245/Benthic%20Density%20Data.csv")

# filter out just juvenile kelp (species = 2002.5)
 jkelpKFM <- jkelpKFM %>% filter(Species ==2002.5)

 
```

## swell height data

import the swell height data that was formatted in "format_data.Rmd"

```{r}
# import swell height data (mean swell height across all Landsat pixels intersecting each transect in each quarter between 1994 and 2009)
lter_waves <- read.csv("../../../intermediate_data_output/GLMM_data/lter_waves.csv")

kfm_waves <- read.csv("../../../intermediate_data_output/GLMM_data/kfm_waves.csv")


```

# format data

```{r}

# plant biomass
# LTER plant biomass data
lter_kelp2 <- lter_kelp %>% mutate(site.trans = paste(SITE, TRANSECT, sep = ".")) %>% rename(year = YEAR) %>% filter(site.trans %in% full_datalvals$site.trans) %>% filter(year %in% full_datalvals$year) %>% select(-TAXON_KINGDOM, -TAXON_PHYLUM, - TAXON_CLASS, -TAXON_ORDER, -TAXON_FAMILY, -TAXON_GENUS, -GROUP, -SURVEY, -MOBILITY, -GROWTH_MORPH) %>% left_join(kelp_regpars, by = "MONTH") %>%  mutate(BIOMASS = (FRONDS*Slope + Intercept)/0.094)%>% mutate(PLANTS = if_else(FRONDS>0, 1, if_else(FRONDS==0, 0, as.double(NA)))) %>% group_by(year, MONTH, DATE, site.trans, QUAD, SIDE, AREA) %>% summarize(n_plants = sum(PLANTS, na.rm=T), AREA = max(AREA),mean_mass = mean(BIOMASS, na.rm = T)) %>% group_by(year, site.trans) %>% summarize(kelp_dens = sum(n_plants)/sum(AREA), mean_mass = mean(mean_mass))%>% filter(kelp_dens>0 & is.na(mean_mass)==F) %>% dplyr::select(-kelp_dens) #%>% filter(mean_mass>0)

# KFM plant biomass data
kfm_kelp2 <- kfm_kelp  %>% mutate(Site = as.character(Site)) %>% filter(Site %in% full_datalvals$site_name) %>% filter(Year %in% full_datalvals$year) %>% rename(site_name = Site) %>% rename(year = Year)
# convert date to date format in order to get the month
kfm_kelp2$Date <- as.Date(kfm_kelp2$Date,format='%d%b%Y')
kfm_kelp2 <- kfm_kelp2 %>% mutate(MONTH = month(Date))%>% left_join(kelp_regpars, by = "MONTH") %>%  mutate(BIOMASS = (StipeCount*Slope + Intercept)/0.094) %>% group_by(site_name, year) %>% summarize(mean_mass=mean(BIOMASS))

# juvenile kelp densities
# LTER data
jkelpLTER2 <- jkelpLTER %>% group_by(YEAR, MONTH, DATE, SITE, TRANSECT, QUAD, SIDE, SP_CODE, COMMON_NAME) %>% summarize(AREA=max(AREA), DENS = sum(COUNT)/AREA) %>% mutate(site.trans = paste(SITE, TRANSECT, sep = ".")) %>% rename(year = YEAR) %>% filter(site.trans %in% full_datalvals$site.trans) %>% filter(year %in% full_datalvals$year) %>% group_by(site.trans, year) %>% summarize(mean_jdens =mean(DENS)) 

# KFM data
jkelpKFM2 <- jkelpKFM %>% mutate(Site = as.character(Site)) %>% filter(Site %in% full_datalvals$site_name) %>% filter(Year %in% full_datalvals$year) %>% filter(is.na(Date)==F) %>% rename(mean_jdens = DensityMean) %>% rename(site_name = Site) %>% rename(year= Year) 


# waves
wave.thresh <- 2.5# threshold for big storm (from Bell et al. 2015)

# turn wave height into binary indicator of whether there was a severe storm
lter_waves2 <- lter_waves %>% filter(year <= max(full_datalvals$year)) %>% filter(year >= min(full_datalvals$year)) %>% group_by(year, site.trans) %>% summarize(waveht = max(waveht)) %>% mutate(Istorm = if_else(waveht>=wave.thresh, 1, 0))%>% ungroup()  

# get severe storm frequency in each transect
lter_waves3 <- lter_waves2 %>% group_by(site.trans) %>% mutate(n_obs = n()) %>% ungroup() %>% group_by(site.trans, n_obs) %>% summarise(n_storms = sum(Istorm)) %>% mutate(storm_freq = n_storms/n_obs) %>% filter(site.trans %in% full_datalvals$site.trans)

kfm_waves2 <- kfm_waves %>% filter(year <= max(full_datalvals$year)) %>% filter(year >= min(full_datalvals$year)) %>% group_by(year, Site) %>% summarize(waveht = max(waveht)) %>% mutate(Istorm = if_else(waveht>=wave.thresh, 1, 0))%>% ungroup()  %>% rename(site_name = Site) %>% mutate(site_name = as.character(site_name)) 

kfm_waves3 <- kfm_waves2 %>% group_by(site_name) %>% mutate(n_obs = n()) %>% ungroup() %>% group_by(site_name, n_obs) %>% summarise(n_storms = sum(Istorm)) %>% mutate(storm_freq = n_storms/n_obs) %>% filter(site_name %in% full_datalvals$site_name)

#View(kfm_waves3)

# combine all variables
# before combining, adjust the wave events to the "kelp year" (e.g., if storm occurred in Nov 2005, it would affect kelp measured in aug 2006 if at all, so would want to change the corresponding year to 2006 to match with the juvenile kelp densities)
#View(lter_waves) # need to get the max wave height each year, filter out just the rows when waveht = max, then if quarter = 3 or 4, change year to the next year. Also if quarter = 1 or 2, need to apply that to quarters 3 and 4 of that same year

lter_waves2.2 <- lter_waves %>% filter(year <= max(full_datalvals$year)) %>% filter(year >= min(full_datalvals$year))%>% mutate(year= if_else(quarter %in% c(3, 4), year + 1, as.numeric(year))) %>% group_by(year, site.trans) %>% mutate(wavehtm = max(waveht)) %>% mutate(Istorm2 = if_else(waveht>=wave.thresh, 1, 0))%>% ungroup() %>% filter(waveht == wavehtm) %>% filter(site.trans %in% full_datalvals$site.trans) %>% dplyr::select(site.trans, year, Istorm2) #mutate(year2 = if_else(quarter %in% c(1, 2), year - 1, as.numeric(NA))) %>% gather(year, year:year2, factor_key=FALSE) %>% filter(is.na(year)==F) %>% distinct()

kfm_waves2.2 <- kfm_waves %>% filter(year <= max(full_datalvals$year)) %>% filter(year >= min(full_datalvals$year))%>% mutate(year= if_else(quarter %in% c(3, 4), year + 1, as.numeric(year)))  %>% group_by(year, Site) %>% mutate(wavehtm = max(waveht)) %>% mutate(Istorm2 = if_else(waveht>=wave.thresh, 1, 0))%>% ungroup() %>% filter(waveht == wavehtm)%>% rename(site_name = Site) %>% mutate(site_name = as.character(site_name))  %>% filter(site_name %in% full_datalvals$site_name) %>% dplyr::select(site_name, year, Istorm2)

#View(lter_waves2.2)
#View(lter_waves2)

# join all the different random effect parameter datasets together for the LTER and KFM transects
# lter
LTERrand <- left_join(full_datalvals, jkelpLTER2, by = c("year", "site.trans")) %>% select(urchin_dens, kelp_dens, year, site.trans, mean_jdens, site_name) %>% filter(is.na(mean_jdens)==F) %>% left_join(lter_kelp2, by = c("year", "site.trans")) %>% left_join(lter_waves2, by = c("year", "site.trans")) %>% left_join(lter_waves2.2, by = c("year", "site.trans"))

# kfm
KFMrand <- left_join(full_datalvals, jkelpKFM2, by = c("year", "site_name")) %>% select(urchin_dens, kelp_dens, year, site_name, mean_jdens, site.trans)%>% filter(is.na(mean_jdens)==F) %>% left_join(kfm_kelp2, by = c("year", "site_name")) %>% left_join(kfm_waves2, by = c("year", "site_name")) %>% left_join(kfm_waves2.2, by = c("year", "site_name"))

# combine lter and kfm data together, also add lag in adult kelp density (since these are the initial densities in the ODE simulations)
allrand <- rbind(LTERrand, KFMrand) %>% group_by(site.trans) %>% mutate(kelp_densLag =lag(kelp_dens,n=1L, default = NA, order_by = year), j_densLag = lag(mean_jdens,n=1L, default = NA, order_by = year)) %>% ungroup() 

#View(allrand)

```

# calculate random effect pars

calculate the values and frequencies of all the random effect parameters (Table S5)

```{r}
# plant biomass
# join the lter and kfm data together
kelp_mass <- c(lter_kelp2$mean_mass, kfm_kelp2$mean_mass)

# get the quintiles of kelp biomass
b.int <- unname(c(quantile(kelp_mass, na.rm = T, 0), quantile(kelp_mass, na.rm = T, 0.2), quantile(kelp_mass, na.rm = T, 0.4), quantile(kelp_mass, na.rm = T, 0.6), quantile(kelp_mass, na.rm = T, 0.8), quantile(kelp_mass, na.rm = T, 1)))

# get the means of each quintile
b.vals <- c(mean(kelp_mass[which(kelp_mass<b.int[2])]), mean(kelp_mass[which(kelp_mass>=b.int[2] &kelp_mass<b.int[3])]), mean(kelp_mass[which(kelp_mass>=b.int[3] &kelp_mass<b.int[4])]), mean(kelp_mass[which(kelp_mass>=b.int[4] &kelp_mass<b.int[5])]), mean(kelp_mass[which(kelp_mass>=b.int[5])]))

b.vals #1.783148,  4.158975,  7.394042, 12.845027, 34.354670

#hist(kelp_mass, n = 100, main = "kelp biomass data", xlab = "mean kg per individual")
#abline(v = b.int[1], lwd = 1, lty=1)
#abline(v = b.int[2], lwd = 1, lty=1)
#abline(v = b.int[3], lwd = 1, lty=1)
#abline(v = b.int[4], lwd = 1, lty=1)
#abline(v = b.int[5], lwd = 1, lty=1)
#abline(v = b.int[6], lwd = 1, lty=1)
#lines(x = b.vals, y = rep(0, length(b.vals)), type = "p", col="red", pch = 16)
#points(x = b.vals[1], y = 0, type = "p", col=b.cols[1], pch = 16, cex = 1.2)
#points(x = b.vals[2], y = 0, type = "p", col=b.cols[2], pch = 16, cex = 1.2)
#points(x = b.vals[3], y = 0, type = "p", col=b.cols[3], pch = 16, cex = 1.2)
#points(x = b.vals[4], y = 0, type = "p", col=b.cols[4], pch = 16, cex = 1.2)
#points(x = b.vals[5], y = 0, type = "p", col=b.cols[5], pch = 16, cex = 1.2)

# juvenile densities
# join the lter and kfm data
j_dens <- c(jkelpLTER2$mean_jdens, jkelpKFM2$mean_jdens)

hist(j_dens, n = 50)

hist(j_dens[which(j_dens>0)], n = 50)
abline(v=quantile(j_dens[which(j_dens>0)], 0.8))

# 80% quantile of non-zero densities
quantile(j_dens[which(j_dens>0)], 0.8)

mean(j_dens[which(j_dens >0 &j_dens <= quantile(j_dens[which(j_dens>0)], 0.8))]) #1.3, occurs 80% of 60% = 0.48

mean(j_dens[which(j_dens > quantile(j_dens[which(j_dens>0)], 0.8))]) # 8.3, occurs 20% of 60% = 0.12

# fraction that are 0
length(which(j_dens==0))/length(j_dens)# about 40%



# storm disturbance
mean(c(kfm_waves3$storm_freq, lter_waves3$storm_freq)) # 0.224, round up to 0.25 since there could be other sources of disturbance other than large swells


```

based on the above, choose a set of plant biomass values of b.vals = c(1.783148,  4.158975,  7.394042, 12.845027, 34.354670) each with prob of 0.2; poor, intermediate, and favorable recruitment probabilities of 0.4, 0.48, and 0.12 (with initial juvenile sporophyte densities of 0, 1.3, and 8.3, and assume higher initial conditions reflect higher maturation rates and lower mortality rates, i.e., more favorable recruitment conditions); and probability fo a severe storm of 0.25


# supp figs

## Fig. S21

distribution of initial kelp densities

```{r}
# look at the distribution of observed kelp densities
pdf("A0dist.pdf", height = 3.5, width = 8)#width = 12, height = 6
par(mfrow = c(1, 2), oma=c(1,2,1,2))
par(mai=c(0.8,0.8,0.01,0.01))
hist(full_datalvals$kelp_dens, n=50, xlab = NA, main = NA, ylab = NA, cex.lab = 1.2)
mtext(side = 3, "a) Observed kelp densities", adj = 0, cex = 1.2)
mtext(side = 1, expression("Kelp density (individuals/m"^2*")"), line = 2.5, cex = 1)
mtext(side = 2, "Frequency", line = 2.5, cex = 1)
# repeat for only low urchin densities
plot(x = sort(full_datalvals$kelp_dens)[1], y = 1, pch =2, xlim = c(0, max(full_datalvals$kelp_dens)), ylim = c(0, 1), xlab = NA, ylab = NA, cex.lab = 1.2, col = "gray46", cex = 0.5)#col = "aquamarine3",
mtext(side = 3, "b) Cumulative probabilities", adj = 0, cex = 1.2)#Cumulative probabilities of initial kelp densities
mtext(side = 1, expression("Kelp density"[italic(i)]), line = 2.5, cex = 1)
mtext(side = 2, expression("P(kelp">="kelp density"[italic(i)]*")"), line = 2.5, cex = 1)
# subset of data with lower urchin densities
for(i in 1:length(full_datalvals$kelp_dens[which(full_datalvals$urchin_dens<25)])){
  points(x = sort(full_datalvals$kelp_dens[which(full_datalvals$urchin_dens<25)])[i], y = length(full_datalvals$kelp_dens[which(full_datalvals$urchin_dens<25)][which(full_datalvals$kelp_dens>=sort(full_datalvals$kelp_dens[which(full_datalvals$urchin_dens<25)])[i])])/length(full_datalvals$kelp_dens[which(full_datalvals$urchin_dens<25)]), pch = 1, col = "lightblue3", cex = 0.55) #, col = "coral2"
}
# full data
for(i in 2:length(full_datalvals$kelp_dens)){
  points(x = sort(full_datalvals$kelp_dens)[i], y = length(full_datalvals$kelp_dens[which(full_datalvals$kelp_dens>=sort(full_datalvals$kelp_dens)[i])])/length(full_datalvals$kelp_dens), pch = 2, col = "gray46", cex = 0.5) #, col = "aquamarine3"
}
abline(v = 0)
legend(x = "topright", legend = c("Full data", "Low urchins"), col = c("gray35", "lightblue3"), bty = "n", pch = c(2, 1), cex = 1.2)
dev.off()

#quantile(full_datalvals$urchin_dens, 0.75) # 24.2


```

```{r}

# estimate probability as the number of observations greater than the predicted Acrit value
plot(x = sort(full_datalvals$kelp_dens)[1], y = 1, pch = 1, xlim = c(0, max(full_datalvals$kelp_dens)), ylim = c(0, 1), xlab = "kelp density i", ylab = "prob(kelp >= kelp density i)", main = "b) Cumulative probabilities for initial kelp densities")
for(i in 2:length(full_datalvals$kelp_dens)){
  points(x = sort(full_datalvals$kelp_dens)[i], y = length(full_datalvals$kelp_dens[which(full_datalvals$kelp_dens>=sort(full_datalvals$kelp_dens)[i])])/length(full_datalvals$kelp_dens), pch = 1)
}
abline(v = 0)

#length(full_datalvals$kelp_dens[which(full_datalvals$kelp_dens>3)])/length(full_datalvals$kelp_dens)


```

### carrying capacity
Check these kelp densities are consistent with a carrying capacity of 1 plant/m2 (see "ODE_parameterization.Rmd")

```{r}
quantile(full_datalvals$kelp_dens, 0.9) # 1.14 ind/m2

```

## Fig. S22
make all the comparison plots to check for relationships between the different random effects parameters

```{r}

# biomass vs. juvenile density

pdf("Reffects.pdf", width = 9)
par(mfrow = c(2, 2), oma=c(1,1.5,1,2))
par(mai=c(0.8,0.8,0.05,0.01))
plot(x = allrand$mean_mass, y = allrand$mean_jdens, xlab = NA, ylab = NA, las = 1)
mtext(side = 1, "Biomass per plant (kg)", line = 2.5)
mtext(side = 2, expression("Juvenile kelp density (ind./m"^2*")"), line = 2.5)
mtext(side = 3, "a) Plant biomass and juvenile densities", adj = 0)

plot(x = log(allrand$mean_mass+1), y = log(allrand$mean_jdens+1), xlab = NA, ylab = NA, las = 1)
mtext(side = 1, "log(biomass per plant + 1)", line = 2.5)
mtext(side = 2, "log(juvenile kelp density + 1)", line = 2.5)
mtext(side = 3, "b) Plant biomass and juvenile densities, log scale", adj = 0)

# storm vs. biomass
boxplot(mean_mass ~ as.factor(Istorm2), data =allrand, xlab = NA, ylab = NA, xaxt = "n", las = 1)
axis(side =1, at = c(1, 2), labels = c("No storm", "Storm"))
mtext(side = 1, "Storm disturbance", line = 2.5)
mtext(side = 2, "Biomass per plant (kg)", line = 2.5)
mtext(side = 3, "c) Plant biomass and storm disturbance", adj = 0)
text(1, 125, "N=278") # length(allrand$Istorm2[which(allrand$Istorm2==0)])
text(2, 125, "N=71") # length(allrand$Istorm2[which(allrand$Istorm2==1)])

# storm vs. initial juvenile density 
boxplot(mean_jdens ~ as.factor(Istorm2), data =allrand, xlab = NA, ylab = NA, xaxt = "n", las = 1)
axis(side =1, at = c(1, 2), labels = c("No storm", "Storm"))
mtext(side = 1, "Storm disturbance", line = 2.5)
mtext(side = 2, expression("Juvenile kelp density (ind./m"^2*")"), line = 2.5)
mtext(side = 3, "d) Juvenile density and storm disturbance", adj = 0)
text(1, 19, "N=278")
text(2, 19, "N=71")
dev.off()


```



